---
title: 'Head Neck Squamous Cell Carcinoma (HNSC) Exploratory Data Analysis'
author: "Sehyun Oh, Britney Pheng"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
    html:
        fontsize: 14pxs
        toc: true
        top-depth: 3
output: html_document
---
# Initial Setup
## Load packages
```{r packages, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  # BiocManager
  library(GenomicSuperSignature)
  library(curatedTCGAData)
  library(MultiAssayExperiment)
  library(TCGAutils)
  library(ComplexHeatmap)
  
  # CRAN
  library(tidyverse) # includes dplyr, ggplot2, magrittr, tidyr
  library(magick)
  library(wordcloud)
  library(ztable)
  library(metafolio)
})
```

## Create TCGA dataset
```{r}
# data_dir <- "~/Documents/GitHub/GSS/data"
# 
# ## Raw read counts from GSE62944 from ExperimentHub
# tcga <- GSEABenchmarkeR::loadEData("tcga", cache = FALSE, paired = FALSE, map2entrez = FALSE)
# 
# ## log2 transformation
# assay(tcga$COAD) <- log2(assay(tcga$COAD) + 1)
# assay(tcga$HNSC) <- log2(assay(tcga$HNSC) + 1)
# 
# TCGA_validationDatasets <- vector(mode = "list", length = 2)
# names(TCGA_validationDatasets) <- c("COAD", "HNSC")
# TCGA_validationDatasets[[1]] <- tcga$COAD
# TCGA_validationDatasets[[2]] <- tcga$HNSC
# 
# ## TCGA-OVC dataset from curatedOvarianData
# BiocManager::install('curatedOvarianData')
# library(curatedOvarianData)
# data(TCGA.RNASeqV2_eset)
# x <- as(TCGA.RNASeqV2_eset, "SummarizedExperiment")
# 
# rs <- rowSums(assay(x) > 2)
# keep <-  rs >= ncol(x) / 2
# tcga_ovc <- x[keep,]
# TCGA_validationDatasets[["OV"]] <- tcga_ovc
# 
# save(TCGA_validationDatasets, file = file.path(data_dir, "TCGA_validationDatasets.rda"))
```

## Load TCGA dataset
```{r}
load('~/Documents/GitHub/GSS/data/TCGA_validationDatasets.rda')
datasets <- TCGA_validationDatasets[1:3]
```

## Load RAVmodel
```{r message=FALSE, warning=FALSE}
RAVmodel <- getModel('C2', load=TRUE)
```

## Select HNSC RNA metadata
```{r message=FALSE, warning=FALSE}
hnsc <- curatedTCGAData(diseaseCode = 'HNSC',
                        assays = 'RNA*',
                        version = '2.0.1',
                        dry.run = FALSE)

hnsc_rna <- getWithColData(hnsc,
                           'HNSC_RNASeq2Gene-20160128',
                           mode = 'append')

hnsc_meta <- colData(hnsc_rna)
```

# heatmapTable: HNSC
```{r message=FALSE}
validate_hnsc <- validate(datasets[['HNSC']], RAVmodel)
heatmapTable(validate_hnsc, RAVmodel)
```

```{r message=FALSE}
assay(hnsc_rna) <- log2(assay(hnsc_rna) + 1)

validate_hnsc_rna <- validate(hnsc_rna, RAVmodel)
heatmapTable(validate_hnsc_rna, RAVmodel)
```

# Subset
## Filter attributes
```{r}
sparsity_summary <- table(colSums(is.na(hnsc_meta)))
sparsity_summary
```

## Sparsity Plot
```{r echo=FALSE}
plot(stack(sparsity_summary)$ind,
     stack(sparsity_summary)$values)
```
```{r}
# Select columns with >10% completeness
keep_attribute_ind <- which(colSums(!is.na(hnsc_meta)) > round(nrow(hnsc_meta)/10))
meta_sub1 <- hnsc_meta[keep_attribute_ind]
meta_sub1 <- subset(meta_sub1, select= -patientID)
```


```{r}
# Randomly select for 100 rows
set.seed(1)
random_sample_ind <- sample(1:nrow(meta_sub1), 100)
meta_sub2 <- meta_sub1[random_sample_ind,]
```

```{r}
# Check for data types in listData
unique(sapply(hnsc_meta@listData, type))

charcTb <- meta_sub2[, sapply(meta_sub1, class) == 'character']
numTb <- meta_sub2[, sapply(meta_sub1, class) %in% c('numeric', 'integer')]
```

```{r}
# Calculate validation scores
sampleScore <- calculateScore(hnsc_rna, RAVmodel)
```

```{r}
validated_ind <- validatedSignatures(validate_hnsc_rna, num.out = 30, RAVmodel, swCutoff = 0.4, indexOnly = TRUE)

# Subset sampleScore to join with MCPcounter
sampleScore_sub <- sampleScore[random_sample_ind, validated_ind] %>% as.data.frame()
```

# Calculate R-Squared Value for Numeric Variables
```{r calculateRsq}
# R squared value function
calculateRsq <- function (x, y) stats::cor(x, y, use = 'na.or.complete') ^ 2
```

```{r warning=FALSE}
# Calculate r-squared for numeric attributes
rsq_numAttr <- as.data.frame(matrix(nrow = ncol(numTb),
                                    ncol = ncol(sampleScore_sub)))

colnames(rsq_numAttr) <- colnames(sampleScore_sub)
rownames(rsq_numAttr) <- colnames(numTb)

for (i in seq_len(ncol(numTb))) {
  for (j in seq_len(ncol(sampleScore_sub))) {
    rsq <- calculateRsq(numTb[, i], sampleScore_sub[, j])
    rsq_numAttr[i, j] <- rsq
  }
}

rsq_numAttr <- na.omit(rsq_numAttr)
```

```{r}
# Remove batch effect variables
batch_num_ind <- grep('analyte|portion|procurement|aliquot|uuid|barcode',
                  rownames(rsq_numAttr))
rsq_numAttr_no_batch <- rsq_numAttr[-batch_num_ind,]
```

```{r}
max_rav <- apply(rsq_numAttr_no_batch, 1, max)
max_attr <- which(max_rav > 0.3)

target_rsq <- rsq_numAttr_no_batch[max_attr,]
```

# heatmapTable
```{r results='asis'}
options(ztable.type='html')
z = ztable(target_rsq)
z %>% makeHeatmap(palette='Purples')
```

```{r}
heatmap(as.matrix(target_rsq))
```

# Calculate F-statistic (ANOVA) for Character Variables
```{r}
# Convert to factor data type
factorTb <- meta_sub2[, sapply(meta_sub2, class) == 'character']

factorTb[sapply(factorTb, is.character)] <- lapply(factorTb[sapply(factorTb, is.character)], factor, exclude = NULL)

factor_ind <- c()

# Select for factors with at least two possible values
for (i in 1:length(factorTb)) {
  if (nlevels(factorTb[,i]) > 1) {
    factor_ind <- c(factor_ind, i)
  }
}

new_factorTb <- factorTb[,factor_ind]
```

```{r}
aov_res <- as.data.frame(matrix(nrow = ncol(new_factorTb),
                                ncol = ncol(sampleScore_sub)))

rownames(aov_res) <- colnames(new_factorTb)
colnames(aov_res) <- colnames(sampleScore_sub)

aov_hnsc_fvalue <- aov_res
aov_hnsc_pvalue <- aov_res

for (i in seq_len(ncol(sampleScore_sub))) {
  for (j in seq_len(ncol(new_factorTb))) {
    
    ## ANOVA
    aov <- aov(sampleScore_sub[, i] ~ new_factorTb[, j])
    
    ## F-statistic
    fval <- summary(aov)[[1]]$`F value`[1]
    aov_hnsc_fvalue[j, i] <- fval
    
    ## p-value
    pval <- summary(aov)[[1]]$`Pr(>F)`[1]
    aov_hnsc_pvalue[j, i] <- pval
  }
}
```

```{r}
# Select for p-values < 0.01
min_rav <- apply(aov_hnsc_pvalue, 1, min)
min_attr <- which(min_rav < 0.01)

target_hnsc_aov_fvalue <- aov_hnsc_fvalue[min_attr,]
target_hnsc_aov_pvalue <- aov_hnsc_pvalue[min_attr,]
```

```{r}
batch_char_ind <- grep('analyte|portion|procurement|aliquot|uuid|barcode',
                       rownames(target_hnsc_aov_fvalue))
hnsc_aov_fvalue <- target_hnsc_aov_fvalue[-batch_char_ind,]
hnsc_aov_pvalue <- target_hnsc_aov_pvalue[-batch_char_ind,]
```

```{r}
heatmap(as.matrix(hnsc_aov_fvalue))
```

```{r}
sig_fval <- as.data.frame(matrix(nrow = ncol(new_factorTb),
                                ncol = ncol(sampleScore_sub)))

rownames(sig_fval) <- colnames(new_factorTb)
colnames(sig_fval) <- colnames(sampleScore_sub)

for (i in seq_len(ncol(sampleScore_sub))) {
  for (j in seq_len(ncol(new_factorTb))) {
    
    if (!is.null(summary(aov(sampleScore_sub[, i] ~ new_factorTb[, j]))[[1]]$`Pr(>F)`[1])) {
      
      if (!is.null(summary(aov(sampleScore_sub[, i] ~ new_factorTb[, j]))[[1]]$`F value`[1]) &
          (summary(aov(sampleScore_sub[, i] ~ new_factorTb[, j]))[[1]]$`Pr(>F)`[1] < 0.05)) {
        sig_fval[j, i] <- summary(aov(sampleScore_sub[, i] ~ new_factorTb[, j]))[[1]]$`F value`[1]
      } else {
        next
      }
    } else {
      next
    }
  }
}

fstat_ind <- c()

for (i in seq_len(nrow(sig_fval))) {
  if (sum(is.na(sig_fval[i, ])) < 11) {
    fstat_ind <- c(fstat_ind, i)
  } else {
    next
  }
}

sig_fval <- sig_fval[fstat_ind,]

batch_char_ind_2 <- grep('analyte|portion|procurement|aliquot|uuid|barcode',
                  rownames(sig_fval))
sig_fval <- sig_fval[-batch_char_ind_2,]

```

```{r results='asis', message=FALSE}
options(ztable.type='html')
ztable(sig_fval) %>%
  makeHeatmap(palette = 'Blues') %>%
  print(caption='ANOVA F-stats with p-values < 0.01')
```

# Score Plot (Categorical Variable)
```{r}
sampleScore1 <- 1001
sampleScore2 <- 4486

plot_data <- sampleScore_sub[, c("RAV1001", "RAV4486")]
colnames(plot_data)[1] <- "sampleScore1"
colnames(plot_data)[2] <- "sampleScore2"

plot_data <- data.frame(plot_data, new_factorTb)

plot_data_2 <- plot_data[-which(is.na(factor(plot_data$pathology_T_stage))),]
```

```{r}
colors <- gg_color_hue(length(unique(plot_data_2$pathology_T_stage)))
colors.toplot <- c(colors)

pA <- ggplot(plot_data_2,
             aes(x = sampleScore1, y = sampleScore2, color = pathology_T_stage)) +
  geom_point() +
  ggtitle('Score Plot for Pathology M Stage Variable') +
  scale_color_manual(values = colors.toplot, name = "Pathology M Stage") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  xlab(paste0("RAV", sampleScore1)) + ylab(paste0("RAV", sampleScore2))

print(pA)
```
```{r}
sampleScore1 <- 1001
sampleScore2 <- 1002

plot_data <- sampleScore_sub[, c("RAV1001", "RAV1002")]
colnames(plot_data)[1] <- "sampleScore1"
colnames(plot_data)[2] <- "sampleScore2"

plot_data <- data.frame(plot_data, new_factorTb)

colors <- gg_color_hue(length(unique(plot_data$patient.gender
)))
colors.toplot <- c(colors)

pA <- ggplot(plot_data,
             aes(x = sampleScore1, y = sampleScore2, color = patient.gender
)) +
  geom_point() +
  ggtitle('Score Plot for Gender') +
  scale_color_manual(values = colors.toplot, name = "Gender") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  geom_hline(yintercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = 0, linetype = 'dashed') +
  xlab(paste0("RAV", sampleScore1)) + ylab(paste0("RAV", sampleScore2))

print(pA)
```
```{r}
t.test(sampleScore2 ~ patient.gender, data = plot_data)
t.test(sampleScore1 ~ patient.gender, data = plot_data)
```


# Scatter Plot (Continuous Variable)
```{r}
hnsc_df <- hnsc_meta %>% as.data.frame
rownames(hnsc_df) <- NULL
hnsc_df$patientID <- gsub("-", "\\.", hnsc_df$patientID)

hnsc_selection <- hnsc_df %>% select(patientID, patient.days_to_death)

sampleScore_sub_3 <- cbind(patientID = rownames(sampleScore_sub), sampleScore_sub)
rownames(sampleScore_sub_3) <- NULL

ss_count <- left_join(hnsc_selection, sampleScore_sub_3, by = "patientID")
ss_count <- na.omit(ss_count)
```

```{r}
estimate_plot <- ggplot(ss_count, aes(RAV1001, patient.days_to_death)) + geom_point() +
geom_smooth(method = 'lm', se=FALSE) +
ylab('Radiation: Day of Form Completion')

estimate_plot
```

## RAV Exploration
```{r}
ind <- 1001
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
drawWordcloud(RAVmodel, ind)
```

```{r}
ind <- 4113
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
drawWordcloud(RAVmodel, ind)
```


```{r}
paste0('RAV',sampleScore1)
```


```{r}
# Function to Display Categorical Attributes' Score Plots

graph_categorical <- function(a, b, c) {
  sampleScore1 <- paste0('RAV', a)
  sampleScore2 <- paste0('RAV', b)
  
  plot_data <- sampleScore_sub[, c(sampleScore1, sampleScore2)]
  colnames(plot_data)[1] <- "sampleScore1"
  colnames(plot_data)[2] <- "sampleScore2"
  
  plot_data <- data.frame(plot_data, new_factorTb)
  
  colors <- gg_color_hue(length(unique(c)) - 1)
  colors.toplot <- c(colors)

  pA <- ggplot(plot_data,
             aes(x = sampleScore1, y = sampleScore2, color = c
                 )) +
        geom_point() +
        #ggtitle('Score Plot for Gender') +
        #scale_color_manual(values = colors.toplot, name = "Gender") +
        theme_bw() +
        theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
        geom_hline(yintercept = 0, linetype = 'dashed') +
        geom_vline(xintercept = 0, linetype = 'dashed') +
        xlab(sampleScore1) + ylab(sampleScore2)

print(pA)
}

graph_categorical(2673, 998, plot_data$pathology_N_stage)
graph_categorical(998, 1959, plot_data$patient.primary_lymph_node_presentation_assessment)
graph_categorical(1125, 3033, plot_data$patient.drugs.drug.2.therapy_types.therapy_type)
```

