---
title: "TCGA Ext"
author: "Britney Pheng"
date: "`r Sys.Date()`"
output: html_document
---

# Setup
Extend prior GSS-studies on five TCGA datasets (COAD, READ, BRCA, LUAD, UCEC) by applying the RAVmodel on seven additional TCGA datasets (BLCA, HNSC, KICH, KIRC, KIRP, LIHC, LUSC).

# Load Packages
```{r}
suppressPackageStartupMessages({
  library(GenomicSuperSignature)
  library(dplyr)
  library(ggplot2)
  library(magick)
  library(wordcloud)
  library(EnrichmentBrowser)
})
```

# Create New TCGA Dataset
```{r include = FALSE}
dat_dir <- "~/Documents/GitHub/GSS/data"

#Raw read counts from GSE62944 from ExperimentHub
tcga <- GSEABenchmarkeR::loadEData("tcga", cache = FALSE, paired = FALSE, map2entrez = FALSE)

## log2 transformation
assay(tcga$BLCA) <- log2(assay(tcga$BLCA) + 1)
assay(tcga$HNSC) <- log2(assay(tcga$HNSC) + 1)
assay(tcga$KICH) <- log2(assay(tcga$KICH) + 1)
assay(tcga$KIRC) <- log2(assay(tcga$KIRC) + 1)
assay(tcga$KIRP) <- log2(assay(tcga$KIRP) + 1)
assay(tcga$LIHC) <- log2(assay(tcga$LIHC) + 1)
assay(tcga$LUSC) <- log2(assay(tcga$LUSC) + 1)

TCGA_validationDatasets <- vector(mode = "list", length = 7)
names(TCGA_validationDatasets) <- c("BLCA", "HNSC", "KICH", "KIRC", "KIRP", "LIHC", "LUSC")
TCGA_validationDatasets[[1]] <- tcga$BLCA
TCGA_validationDatasets[[2]] <- tcga$HNSC
TCGA_validationDatasets[[3]] <- tcga$KICH
TCGA_validationDatasets[[4]] <- tcga$KIRC
TCGA_validationDatasets[[5]] <- tcga$KIRP
TCGA_validationDatasets[[6]] <- tcga$LIHC
TCGA_validationDatasets[[7]] <- tcga$LUSC

save(TCGA_validationDatasets, file = file.path(dat_dir, "TCGA_validationDatasets.rda"))
```

## TCGA Datasets
```{r}
load("data/TCGA_validationDatasets.rda")
names(TCGA_validationDatasets)
datasets <- TCGA_validationDatasets[1:7]
```

## RAVmodel
```{r}
RAVmodel <- getModel("C2", load = TRUE)
```

# Comparison
## heatmapTable (all)
```{r}
val_all <- GenomicSuperSignature::validate(datasets, RAVmodel)
heatmapTable(val_all, RAVmodel)
```
```{r}
val_blca <- validate(datasets[['BLCA']], RAVmodel)
heatmapTable(val_blca, RAVmodel)
```

```{r}
val_hnsc <- validate(datasets[['HNSC']], RAVmodel)
heatmapTable(val_hnsc, RAVmodel)
```

```{r}
val_kich <- validate(datasets[['KICH']], RAVmodel)
heatmapTable(val_kich, RAVmodel)
```

```{r}
val_kirc <- validate(datasets[['KIRC']], RAVmodel)
heatmapTable(val_kirc, RAVmodel)
```

```{r}
val_kirp <- validate(datasets[['KIRP']], RAVmodel)
heatmapTable(val_kirp, RAVmodel)
```

```{r}
val_lihc <- validate(datasets[['LIHC']], RAVmodel)
heatmapTable(val_lihc, RAVmodel)
```
```{r}
val_lusc <- validate(datasets[['LUSC']], RAVmodel)
heatmapTable(val_lusc, RAVmodel)
```

## MeSH Terms and Associated Studies

### BLCA (Bladder Urothelial Carcinoma)
```{r echo = FALSE, warning = FALSE}
ind <- 468
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```

```{r echo = FALSE, warning = FALSE}
ind <- 683
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```
### HNSC (Head and Neck Squamous Cell Carcinoma)
```{r echo = FALSE, warning = FALSE}
ind <- 185
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```

### KICH (Kidney Chromophobe)
```{r echo = FALSE, warning = FALSE}
ind <- 357
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```

### KIRC (Kidney Renal Clear Cell Carcinoma)
```{r echo = FALSE, warning = FALSE}
ind <- 468
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```

```{r echo = FALSE, warning = FALSE}
ind <- 625
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```

### KIRP (Kidney Renal Papillary Cell Carcinoma)
```{r echo = FALSE, warning = FALSE}
ind <- 1303
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```

### LIHC (Liver Hepatocellular Carcinoma)
```{r echo = FALSE, warning = FALSE}
ind <- 1669
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```
```{r echo = FALSE, warning = FALSE}
ind <- 1166
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```

### LUSC (Lung Squamous Cell Carcinoma)
```{r echo = FALSE, warning = FALSE}
ind <- 814
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```


