---
title: 'Colon Adenocarcinoma (COAD) EDA: categorical variables'
author: "Sehyun Oh, Britney Pheng"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
    html:
        fontsize: 14pxs
        toc: true
        top-depth: 3
output: html_document
---

# Initial Setup
## Load packages
```{r packages, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
    library(tidyverse)
})
```

## Load data
```{r message=FALSE}
## SummarizedExperiment object containing COAD train dataset 
coad_train <- readRDS("data/tcga_coad_train.rds") # 20,501 genes x 261 samples

## Sample scores for the train set (80% of the randomly selected COAD samples)
## Only the top 15 validated RAVs
sampleScore_sub <- read.csv("data/sampleScore_train.csv", row.names = 1) # 261 samples x 15 RAVs

## Training set's metadata: character variables
charTb <- read.csv("data/meta_train_char.csv", row.names = 1) # 261 samples x 182 metadata attributes (char)
```


# Convert to factor
Convert character variables into the `factor` data type
```{r}
factorTb <- charTb
factorTb[sapply(factorTb, is.character)] <- lapply(factorTb[sapply(factorTb, is.character)], factor)
```

Check the factor level
```{r}
nlevel <- sapply(factorTb, function(x) {
    res <- levels(x) %>% na.omit
    return(length(res))
})

factorTb_1_lv <- factorTb[nlevel == 1]
factorTb_2_lv <- factorTb[nlevel == 2]
factorTb_m_lv <- factorTb[nlevel > 2]

# single_fac_ind <- c()
# binary_fac_ind <- c()
# multi_fac_ind <- c()
# 
# # Testing factor grouping
# for (i in seq_len(ncol(factorTb))) {
#   if (nlevels(factorTb[,i]) == 1 |
#       (nlevels(factorTb[,i]) == 2 &
#        any(is.na(levels(factorTb[,i]))))
#       ) {
#     single_fac_ind <- c(single_fac_ind, i)
#   } else if (#nlevels(factorTb[,i]) == 3 &
#              #any(is.na(levels(factorTb[,i]))) |
# 
#              (nlevels(factorTb[,i]) == 2 &
#               !any(is.na(levels(factorTb[,i]))))
#           ) {
#     binary_fac_ind <- c(binary_fac_ind, i)
#   } else {
#     multi_fac_ind <- c(multi_fac_ind, i)
#   }
# }
# 
# new_factorTb <- factorTb[,multi_fac_ind]
# binary_factor <- factorTb[,binary_fac_ind]
# binary_factor <- binary_factor[-c(7,8)]
# single_factor <- factorTb[,single_fac_ind]
```


# F-statistic (ANOVA) for multi-level variables
This section goes through the following steps:     
    1. Calculate F-statistics between sample scores and categorical variables    
    2. Filter by p-value cutoff    
    3. Plot heatmap of the F-stats results with significant p-value
    
## Calculate F-statistic (ANOVA) 
```{r}
aov_res <- as.data.frame(matrix(nrow = ncol(factorTb_m_lv),
                                ncol = ncol(sampleScore_sub)))

rownames(aov_res) <- colnames(factorTb_m_lv)
colnames(aov_res) <- colnames(sampleScore_sub)

aov_fvalue <- aov_res
aov_pvalue <- aov_res

for (i in seq_len(ncol(sampleScore_sub))) {
  for (j in seq_len(ncol(factorTb_m_lv))) {
    
    ## ANOVA
    aov <- aov(sampleScore_sub[, i] ~ factorTb_m_lv[, j])
    
    ## F-statistic
    fval <- summary(aov)[[1]]$`F value`[1]
    aov_fvalue[j, i] <- fval
    
    ## p-value
    pval <- summary(aov)[[1]]$`Pr(>F)`[1]
    aov_pvalue[j, i] <- pval
  }
}
```

## Filter by p-value cutoff
```{r}
# Select for p-values < 0.01
min_rav <- apply(aov_pvalue, 1, min)
min_attr <- which(min_rav < 0.01)

aov_coad_fvalue <- aov_fvalue[min_attr,]
aov_coad_pvalue <- aov_pvalue[min_attr,]
```


## Heatmap
F-statistics results of multi-level attributes with p-value above 0.01.
```{r}
heatmap(as.matrix(aov_coad_fvalue), main = 'COAD F-Statistics')

png("data/F_statistics_multi_level.png", width = 20, height = 20, units = 'in', res = 600)
heatmap(as.matrix(aov_coad_fvalue), main = 'COAD F-Statistics')
dev.off()
```

```{r}
## p-value cutoff
aov_p_df <- aov_coad_pvalue < 0.05 

## Maximum test value for attributes and RAVs --> We can switch to top `n`
aov_df <- aov_res
aov_df[is.na(aov_df)] <- 0 # Assign 0 for `FALSE`

n = 1  # the number of top t-values to collect. If it is 1, only the maximum t-statics value will be used
row_ind <- apply(aov_coad_fvalue, 1, 
                 function(x) {which(x >= sort(x, decreasing = TRUE)[n])})
col_ind <- apply(aov_coad_fvalue, 2,
                 function(x) {which(x >= sort(x, decreasing = TRUE)[n])})

for (i in seq_len(length(row_ind)/n)) {
    for (k in seq_len(n)) {
        j <- row_ind[i][k]
        if (isTRUE(aov_p_df[i,j])) { # Significant p-value
            aov_df[i, j] <- 1 # Assign 1 for `TRUE`
        }
    }
} 

for (i in seq_len(length(col_ind))) {
    m <- col_ind[[i]] # the number of attributes having the top `n` t-values
    for (k in seq_along(m)) {
        j <- m[k]
        if (isTRUE(aov_p_df[j, i])) {
        aov_df[j, i] <- 1 # Assign 1 for `TRUE`
        }
    }
} 
```

```{r}
aov_df_sig <- aov_df[rowSums(aov_df) != 0,] # high F-statistics values passing the p-value cutoff
heatmap(as.matrix(aov_df_sig))
```


```{r}
## Save the sample score of the multi-variables with high f-test values above the p-value cutoff
write.csv(aov_df_sig, "data/train_char_ftest_p_point05.csv")
write.csv(factorTb_m_lv, "data/meta_train_char_multi.csv")
```

```{r what_is_this_for?, echo=FALSE, eval=FALSE}
# sig_fval <- as.data.frame(matrix(nrow = ncol(factorTb_m_lv),
#                                  ncol = ncol(sampleScore_sub)))
# 
# rownames(sig_fval) <- colnames(factorTb_m_lv)
# colnames(sig_fval) <- colnames(sampleScore_sub)
# 
# for (i in seq_len(ncol(sampleScore_sub))) {
#   for (j in seq_len(ncol(factorTb_m_lv))) {
#     
#     if (!is.null(summary(aov(sampleScore_sub[, i] ~ factorTb_m_lv[, j]))[[1]]$`Pr(>F)`[1])) {
#       
#       if (!is.null(summary(aov(sampleScore_sub[, i] ~ factorTb_m_lv[, j]))[[1]]$`F value`[1]) &
#           (summary(aov(sampleScore_sub[, i] ~ factorTb_m_lv[, j]))[[1]]$`Pr(>F)`[1] < 0.05)) {
#         sig_fval[j, i] <- summary(aov(sampleScore_sub[, i] ~ factorTb_m_lv[, j]))[[1]]$`F value`[1]
#       } else {
#         next
#       }
#     } else {
#       next
#     }
#   }
# }
# 
# fstat_ind <- c()
# 
# for (i in seq_len(nrow(sig_fval))) {
#   if (sum(is.na(sig_fval[i, ])) < 220) {
#     fstat_ind <- c(fstat_ind, i)
#   } else {
#     next
#   }
# }
# 
# sig_fval <- sig_fval[fstat_ind,]
# batch_char_ind_2 <- grep('analyte|portion|procurement|aliquot|uuid|barcode',
#                   rownames(sig_fval))
# sig_fval <- sig_fval[-batch_char_ind_2,]
```


# Calculate t-test for binary variables

## Remove attributes without enough observations
```{r}
too_few_obs_ind <- c()

for (i in seq_len(ncol(factorTb_2_lv))) {
    x <- summary(factorTb_2_lv[,i])
  
    if (all(x > 1)) {
        # print(x)
    } else {
        too_few_obs_ind <- c(too_few_obs_ind, i)
        attr_name <- colnames(factorTb_2_lv)[i]
        # print(paste(attr_name, "does not have enough observations"))
    }
}

binary_factors <- factorTb_2_lv[,-too_few_obs_ind]
```

```{r}
## Save the metadata table for binary variables converted into factors
write.csv(binary_factors, "data/meta_train_char_binary.csv")
```


```{r}
t_test_res <- as.data.frame(matrix(nrow = ncol(binary_factors),
                                   ncol = ncol(sampleScore_sub)))

rownames(t_test_res) <- colnames(binary_factors)
colnames(t_test_res) <- colnames(sampleScore_sub)

ttest_coad_tvalue <- t_test_res
ttest_coad_pvalue <- t_test_res

for (i in seq_len(ncol(sampleScore_sub))) {
    for (j in seq_len(ncol(binary_factors))) {
        ## t-test
        t_test <- t.test(sampleScore_sub[, i] ~ binary_factors[,j])
    
        ## t-value
        tval <- t_test$statistic
        ttest_coad_tvalue[j, i] <- tval
    
        ## p-value
        pval <- t_test$p.val
        ttest_coad_pvalue[j, i] <- pval
    }
}
```

## Filter by p-value cutoff
Filter the attributes associated with the RAV(s) that pass the p-value cutoff
```{r}
# Select for p-values < 0.05
min_rav <- apply(ttest_coad_pvalue, 1, min)
min_attr <- which(min_rav < 0.05)

ttest_coad_tvalue_sig <- ttest_coad_tvalue[min_attr,]
ttest_coad_pvalue_sig <- ttest_coad_pvalue[min_attr,]
```

```{r}
heatmap(as.matrix(ttest_coad_tvalue_sig), main = 'COAD T-Test')

png("data/T_test_binary.png", width = 20, height = 20, units = 'in', res = 600)
heatmap(as.matrix(ttest_coad_tvalue_sig), main = 'COAD T-Test')
dev.off()
```

Subset t-test results with p-value cutoff
```{r}
## p-value cutoff
ttest_p_df <- ttest_coad_pvalue < 0.05 

## Maximum t-test value for attributes and RAVs --> We can switch to top `n`
t_df <- t_test_res
t_df[is.na(t_df)] <- 0 # Assign 0 for `FALSE`

n = 1  # the number of top t-values to collect. If it is 1, only the maximum t-statics value will be used
row_ind <- apply(ttest_coad_tvalue, 1, 
                 function(x) {which(x >= sort(x, decreasing = TRUE)[n])})
col_ind <- apply(ttest_coad_tvalue, 2,
                 function(x) {which(x >= sort(x, decreasing = TRUE)[n])})

for (i in seq_len(length(row_ind)/n)) {
    for (k in seq_len(n)) {
        j <- row_ind[i][k]
        if (isTRUE(ttest_p_df[i,j])) { # Significant p-value
            t_df[i, j] <- 1 # Assign 1 for `TRUE`
        }
    }
} 

for (i in seq_len(length(col_ind))) {
    m <- col_ind[[i]] # the number of attributes having the top `n` t-values
    for (k in seq_along(m)) {
        j <- m[k]
        if (isTRUE(ttest_p_df[j, i])) {
        t_df[j, i] <- 1 # Assign 1 for `TRUE`
        }
    }
} 
```

```{r}
t_df_sig <- t_df[rowSums(t_df) != 0,] # high t-test values passing the p-value cutoff
heatmap(as.matrix(t_df_sig))
```

```{r}
## Save the sample score of the binary variables with high t-test values above the p-value cutoff
write.csv(t_df_sig, "data/train_char_ttest_p_point05.csv")
```

```{r}
t_df_sig %>% filter(RAV834 == 1) %>% select(RAV834) %>% rownames()
```

```{r}
t_df_sig %>% filter(RAV1008 == 1) %>% select(RAV1008) %>% rownames()
```

```{r}
t_df_sig %>% filter(RAV1302 == 1) %>% select(RAV1302) %>% rownames()
```




```{r echo=FALSE, eval=FALSE}
# # Only view t-values with significant p-value (<0.05)
# ttest_coad_sig_tvalue <- t_test_res
# 
#  for (i in seq_len(ncol(sampleScore_sub))) {
#    for (j in seq_len(ncol(binary_factors))) {
#      if (t.test(sampleScore_sub[, i] ~ binary_factors[,j])$p.val < 0.05) {
#        ttest_coad_sig_tvalue[j, i] <- t.test(sampleScore_sub[, i] ~ binary_factors[,j])$statistic
#      }
#    }
#  }
```

```{r results='asis', message=FALSE}
options(ztable.type='html')
ztable(t_df_sig) %>%
  makeHeatmap(palette = 'Blues') %>%
  print(caption = 'T-test with p-values < 0.05')
```
