---
title: 'curatedTCGA metadata: KICH'
author: "Britney Pheng"
date: "`r Sys.Date()`"
output: html_document
---
Kidney Chromophobe (KICH)

# Load Packages
```{r load packages, warning=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(GenomicSuperSignature)
  library(curatedTCGAData)
  library(MultiAssayExperiment)
  library(TCGAutils)
  library(tidyr)
  library(dplyr)
  library(ggplot2)
  library(magick)
  library(wordcloud)
  library(EnrichmentBrowser)
})
```

```{r RAVmodel, echo=FALSE, warning=FALSE, message=FALSE}
RAVmodel <- getModel("C2", load=TRUE)
```

# TCGA Datasets
```{r load datasets}
load("~/Documents/GitHub/GSS/data/TCGA_validationDatasets.rda")
datasets <- TCGA_validationDatasets[1:7]
#head(datasets)
```

# Overview
## heatmapTable (all)
```{r datasets heatmap table}
## heatmapTable
val_all <- validate(datasets, RAVmodel)
heatmapTable(val_all, RAVmodel, scoreCutoff = 0.7)
```
## heatmapTable: KICH
```{r}
val_kich <- validate(datasets[['KICH']], RAVmodel)
heatmapTable(val_kich, RAVmodel)
```
# curated TCGA Data for KICH

## Select table of KICH RNA metadata
```{r MetaData, message=FALSE}
tcga_kich <- curatedTCGAData(diseaseCode = 'KICH', assays = "RNA*", version = "2.0.1", dry.run = FALSE)
head(tcga_kich)
```
## Select for 'KICH_RNASeq2Gene-20160128' metadata with dimensions (423 x 3)
```{r Select SummarizedExperiment assay, warning=FALSE}
(tcga_kich_rna <- getWithColData(tcga_kich, "KICH_RNASeq2Gene-20160128", mode = "append"))
```

## heatmapTable (curatedTCGAData - KICH)
```{r tcga, message = FALSE}
val_tcga_rna <- validate(tcga_kich_rna, RAVmodel)
heatmapTable(val_tcga_rna, RAVmodel)
```

Compared to the prior heatmap tables, RAV 40 now has the highest validation score and average SW. Let's view the studies and biological pathways within the RAV.

## RAV 40
```{r echo = FALSE, warning = FALSE}
ind <- 40
findStudiesInCluster(RAVmodel, ind, studyTitle = TRUE)
subsetEnrichedPathways(RAVmodel, ind, include_nes = TRUE) %>% as.data.frame
drawWordcloud(RAVmodel, ind)
```


